Experience-Biased Symbolic Planning

### 2014-05-12 AAAI_AuthorKit/paper.tex
New section: Analysis of HSPr

\textit{For sufficiently large $\epsilon^E$, once the A* frontier touches an E-graph component which includes a goal state, no node outside the E-graph will ever again need to be expanded.} This means, for instance, that if the E-graph already contains a path from the start to a goal, then A*, guided by $h^E$ with large bias parameter $\epsilon^E$, will never leave the E-graph. If the E-graph is small, a solution would promptly be discovered.

@@

We remark that, when E-graphs are involved, the values $g_S(S')$ on the right-hand side range over a variety of different states $S'$. Compare this to the formula for $h^{HSP}$, which uses only $g_S(G)$. HSP would still compute all the $g_S$-values despite using them only once, so here we get to consult the E-graph essentially for free. The only additional work is the minimization over $S'\in V^E$, which is negligible for small E-graphs. A pseudocode implementation is listed in Algorithms \ref{alg:ComputeG} and \ref{alg:Search}. The parameter $\epsilon^E$ controls bias toward the E-graph, while $\epsilon$ provides an additional goal-directed bias.

\begin{algorithm}
\caption{ComputeG(S)}
\label{alg:ComputeG}
\begin{algorithmic}
\FORALL{$a \in A$}
\IF{$a \in S$}
\STATE $g_S(a) \leftarrow 0$
\ELSE
\STATE $g_S(a) \leftarrow \infty$
\ENDIF
\ENDFOR
\REPEAT
\FORALL{$op \in O$}
\STATE $pCost \leftarrow g_S(Prec(op))$
\FORALL{$a \in Add(op)$}
\STATE $g_S(a) \leftarrow \min \left(g_S(a),~pCost + Cost(op)\right)$
\ENDFOR
\ENDFOR
\UNTIL{no further change in $g$-values}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Search()}
\label{alg:Search}
\begin{algorithmic}
\STATE Let $c(S,S')$ for $S,S'\in V^E$ represent E-graph edge costs.
\FORALL{$S \in V^E$}
\STATE ComputeG(S)
\FORALL{$S' \in V^E$}
\STATE $c'(S,S') \leftarrow \min\left(c(S,S'),~\epsilon^E g_S(S')\right)$
\ENDFOR
\ENDFOR
\STATE Compute $h^E$ on $V^E$ by reverse Dijkstra from $G$ with $c'$.
\STATE Run A* on the full graph from $I$ with heuristic $\epsilon h^E$:
\FOR{$S \notin V^E$ when generated by A*}
\STATE ComputeG(S)
\STATE $h^E(S) \leftarrow \min_{S'\in V^E} \left( \epsilon^E g_S(S') + h^E(S') \right)$
\ENDFOR
\IF{A* successfully found a path to some goal state $S \supseteq G$}
\STATE Add all edges of the solution path to the E-graph.
\RETURN solution path
\ELSE
\RETURN failure
\ENDIF
\end{algorithmic}
\end{algorithm}

ComputeG implements the dynamic programming computation of $g$-values.
Each iteration of the outermost loop permanently fixes at least one $g_S(a)$ value, so it does at most $|A|+1$ iterations.
Multiplying everything together and ignoring cardinality signs for brevity, the runtime bound is $O(\mathbf{A^2O})$.
Thus, each state generated costs $O(\mathbf{A^2O + V^E})$ time.
The preprocessing cost is $O(\mathbf{A^2OV^E + (V^E)^2})$, asymptotically equivalent to preemptively generating every state in $V^E$.
For comparison, standard HSP without E-graphs takes $O(\mathbf{A^2O})$ time (with essentially the same constant factors) to generate a state, and does not incur the cost of generating a state unless the A* encounters it.

@@

\section{Conclusions and Extensions}

>>>>

\textit{For sufficiently large $\epsilon^E$, once the A* frontier touches an E-graph component which includes a goal state, no node outside the E-graph will ever again need to be expanded.}

This means, for instance, that if the E-graph already contains a path from the start to a goal, then A*, guided by $h^E$ with large bias parameter $\epsilon^E$, will never leave the E-graph.
If the E-graph is small, a solution would promptly be discovered.
In this manner, little work is needed to solve previously encountered subproblems.

@@

We remark that, when E-graphs are involved, the values $g_S(S')$ on the right-hand side range over a variety of different states $S'$. Compare this to the formula for $h^{HSP}$, which uses only $g_S(G)$. HSP would still compute all the $g_S$-values despite using them only once, so here we get to consult the E-graph essentially for free. The only additional work is the minimization over $S'\in V^E$, which takes O($|V^E||A|$) time. A pseudocode implementation is listed in Algorithms \ref{alg:ComputeG} and \ref{alg:Search}. The parameter $\epsilon^E$ controls bias toward the E-graph, while $\epsilon$ provides an additional goal-directed bias.

\begin{algorithm}
\caption{ComputeG($S$)}
\label{alg:ComputeG}
\begin{algorithmic}
\FORALL{$a \in A$}
\IF{$a \in S$}
\STATE $g_S(a) \leftarrow 0$
\ELSE
\STATE $g_S(a) \leftarrow \infty$
\ENDIF
\ENDFOR
\REPEAT
\FORALL{$op \in O$}
\STATE $pCost \leftarrow g_S(Prec(op))$
\FORALL{$a \in Add(op)$}
\STATE $g_S(a) \leftarrow \min \left(g_S(a),~pCost + Cost(op)\right)$
\ENDFOR
\ENDFOR
\UNTIL{no further change in $g$-values}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Search()}
\label{alg:Search}
\begin{algorithmic}
\STATE Let $c(S,S')$ for $S,S'\in V^E$ represent E-graph edge costs.
\FORALL{$S \in V^E$}
\STATE ComputeG($S$)
\FORALL{$S' \in V^E$}
\STATE $c'(S,S') \leftarrow \min\left(c(S,S'),~\epsilon^E g_S(S')\right)$
\ENDFOR
\ENDFOR
\STATE Compute $h^E$ on $V^E$ by reverse Dijkstra from $G$ with $c'$.
\STATE Run A* on the full graph from $I$ with heuristic $\epsilon h^E$:
\FOR{$S \notin V^E$ when generated by A*}
\STATE ComputeG($S$)
\STATE $h^E(S) \leftarrow \min_{S'\in V^E} \left( \epsilon^E g_S(S') + h^E(S') \right)$
\ENDFOR
\IF{A* successfully found a path to some goal state $S \supseteq G$}
\STATE Add all edges of the solution path to the E-graph.
\RETURN solution path
\ELSE
\RETURN failure
\ENDIF
\end{algorithmic}
\end{algorithm}

ComputeG implements the dynamic programming computation of $g$-values.
Each iteration of the outermost loop permanently fixes at least one $g_S(a)$ value, so it does at most $|A|+1$ iterations.
Multiplying everything together and ignoring cardinality signs for brevity, the runtime bound is $O(\mathbf{A^2O})$.
Thus, each state generated costs $O(\mathbf{A^2O + V^EA})$ time.
The preprocessing cost is $O(\mathbf{V^EA^2O + (V^E)^2A})$, asymptotically equivalent to preemptively generating every state in $V^E$.
For comparison, standard HSP without E-graphs takes $O(\mathbf{A^2O})$ time (with essentially the same constant factors) to generate a state, and does not incur the cost of generating a state unless the A* encounters it. Provided the number of E-graph vertices is limited to $O(\mathbf{AO})$, the asymptotic cost of node generation is unchanged.

@@

\section{Analysis of HSPr}

Whether or not E-graphs are used, the call to ComputeG($S$) at each node generation is quite substantial. For this reason, a variant of HSP, called HSPr, searches the regression space of STRIPS, where each node represents a ``subgoal". The search proceeds backward from the goal. In this case, the heuristic must estimate the cost to achieve a given state $S$ (considered as a subgoal) from the initial state $I$. A single preprocessing call to ComputeG($I$) allows each $h^{HSPr}(S) = g_I(S)$ to be computed in $O(\mathbf{A})$ time.

With E-graphs, in backward as well as forward search, we call ComputeG($S$) for all $S\in V^E$; only now, $V^E$ includes $I$ instead of $G$. From $I$, a forward Dijkstra yields $h^E(S)$ for all $S\in V^E$. Then for $S\notin V^E$, the E-graph heuristic estimate of the cost from $I$ to $S$ is given by
\[h^E(S) = \min_{S'\in V^E}\left(h^E(S') + \epsilon^E g_{S'}(S) \right)\]

Since $g_{S'}$ was precomputed, the cost of generating a state is reduced to just $O(\mathbf{V^EA})$. Nonetheless, the overhead of using E-graphs is much more substantial in backward than in forward search, contributing a $|V^E|$ factor on the state generation cost.

Since a major benefit of backward HSP is being able to generate states in $O(\mathbf{A})$ time, it appears E-graphs in their present form are less likely to be of much help. However, in domains where forward search is more practical, we saw that E-graphs can effectively provide free hints.

\section{Conclusions}
